---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# maternalhealthtools

<!-- badges: start -->
<!-- badges: end -->

The goal of maternalhealthtools is to provide reusable functions for cleaning, training, testing, and visualizing maternal health data. It includes tools for handling missing or inconsistent data, training and testing multinomial logistic regression and random forest models, and generating plot for exploratory analysis and model interpretation.


## Installation

You can install the development version of maternalhealthtools from:

[GitHub](https://github.com/DSCI-310-2025/maternalhealthtools.git) with:

``` r
# install.packages("devtools")
devtools::install_github("DSCI-310-2025/maternalhealthtools")
```

## Usage

A common task when dealing with maternal health data involves checking for missing values, cleaning data, and examining unique values from the target column.
This is where functions like check_na(), clean(), and get_targets() are useful. 

```{r}
library(maternalhealthtools)

# Sample dataset
df <- data.frame(
  Age = c(25, NA, 35, 40, 45, 30),
  BloodPressure = c(120, 130, NA, 110, 140, 125),
  RiskLevel = c("low risk", "mid risk", "high risk", NA, "high risk", "mid risk"),
  stringsAsFactors = FALSE
)

check_na(df)
```

check_na() summarizes the number of missing values for each column. This makes it easier to spot where cleaning is needed.

```{r}
cleaned_df <- clean(df, RiskLevel)
cleaned_df
```
clean() ensures that the target variable is properly formatted as a factor and removes the missing data.

```{r}
get_targets(cleaned_df, RiskLevel)
```
get_targets() returns the unique values of the target column. This is especially useful when working with large datasets and for understanding the class distribution before modeling.

```{r}
# Example training data
train_data <- data.frame(
  Age = c(25, 30, 35),
  SystolicBP = c(120, 130, 110),
  DiastolicBP = c(80, 85, 75),
  BS = c(5.6, 6.2, 4.9),
  BodyTemp = c(98.6, 99.1, 97.9),
  HeartRate = c(72, 80, 76),
  RiskLevel = factor(c("low risk", "mid risk", "high risk"))
)

mlr_model <- training_mlr_model(train_data)
summary(mlr_model)
```
training_mlr_model() trains a multinomial logistic regression model based on the data. 

```{r}
rf_model <- training_rf_model(train_data)
print(rf_model)
```
Similarly, training_rf_model() builds a random forest model using the default of 500 trees.

The testing() function loads a saved classification model (e.g., .rds file) and applies it to a test dataset to generate predictions. The class probabilities are also available but is disabled by default.

```{r}
# Example usage
# testing("path/to/mlr_model.rds", test_data)
```

Note: Since the function depends on a trained model file, examples are not run in the README file.

The visualization() helps generate three types of plots:

1. Confusion Matrix plot
2. Random Forest Feature Importance plot
3. Predicted Probability vs Blood Sugar plot

This function is useful for model visualization and interpretation. It saves the generated plot to a file in the specified output directory. 

```{r}
# Example usage
# 1. Confusion Matrix Plot example

# helper_conf_table <- data.frame(
#                True = c("low risk", "mid risk", "high risk",
#                        "low risk", "mid risk", "high risk",
#                        "low risk", "mid risk", "high risk"),
#                Predicted = c("low risk", "low risk", "low risk",
#                                "mid risk", "mid risk", "mid risk",
#                                "high risk", "high risk", "high risk"),
#                Frequency = c(9, 17, 9, 10, 6, 9, 8, 14, 18),
#                Percentage = c(33.3, 45.9, 25.0, 37.0, 16.2, 25.0, 29.6, 37.8, 50.0))
#                
# visualization("conf_matrix", helper_conf_table, conf_type = "mlr", output_dir = "outputs")


# 2. Random Forest Feature Importance plot

# rf_model <- training_rf_model(train_data)

# visualization("feature_importance", rf_model, conf_type = NULL, output_dir = "outputs")

# 3. Predicted Probability vs Blood Sugar plot

# prob_table <- data.frame(
#   BS = c(4.5, 5.0, 5.5, 6.0, 6.5, 7.0),
#   Probability = c(0.2, 0.3, 0.5, 0.6, 0.7, 0.9),
#   RiskLevel = factor(c("low risk", "low risk", "mid risk", "mid risk", "high risk", "high risk"))
#   )

# visualization("pred_prob", prob_table, conf_type = NULL, output_dir = "outputs")
```

Note: Since the function depends on input and output files, examples are not run in the README file. For more information on the visualization() function, reading the visualization.R script is recommended.


## Position in the R Package Ecosystem

The maternalhealthtools package provides a small set of helper functions to support data analysis workflows. It combines common tasks such as cleaning data, training and testing models, and generating visualizations. While the functions in this package are built on widely used R packages like dplyr, randomForest, and ggplot2, they are meant to simplify repetitive steps. Compared to other packages, maternalhealthtools provides a lighter and more specific use case like maternal health classification. It offers a predefined workflow that reduces the amount of repeated code. 


