---
title: "Introduction to maternalhealthtools"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to maternalhealthtools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(maternalhealthtools)
library(dplyr)
```

## Introduction

The `maternalhealthtools` package provides a set of tools for cleaning data, training and testing models, and visualizing outcomes for maternal health data. This vignette provides a full example using the `maternalhealthtools` package. 

## 1. Sample Data
We will being by creating a sample data for the example.
```{r}
sample_df <- data.frame(
  Age = c(25, 30, 35, NA, 42, 29, 38, 31, NA, 40),
  SystolicBP = c(120, 130, 135, 115, 140, 125, 135, 128, 122, NA),
  DiastolicBP = c(80, 85, 78, 75, 90, NA, 82, 84, 79, 77),
  BS = c(5.6, 6.2, 4.9, 7.1, NA, 5.8, 6.4, NA, 5.9, 6.3),
  BodyTemp = c(98.6, 99.1, 97.9, 98.4, 98.9, NA, 98.2, 98.7, 98.5, 99.0),
  HeartRate = c(72, 80, 76, 74, 85, 78, NA, 82, 79, 77),
  RiskLevel = c("low risk", "mid risk", "high risk", "low risk", NA,
                "mid risk", "high risk", "low risk", "mid risk", "high risk"),
  stringsAsFactors = FALSE
)

sample_df
```

## 2. Check for Missing Values
We use `check_na()` to identify where missing values exist.
```{r}
check_na(sample_df)
```

## 3. Clean the data
By using `clean()`, we clean the data by removing the rows with missing values and converting the target column `RiskLevel` into a factor.
```{r}
cleaned_df <- clean(sample_df, RiskLevel)
cleaned_df
```

## 4. Get Target Classes
Use `get_targets()` to inspect the unique classes in the target column.
```{r}
get_targets(cleaned_df, RiskLevel)
```

## 5. Train Models
Now that we have clean data, we can start training them. Here is how we train two models, multinomial logistic regression and random forest model, using the provided training functions. 

```{r}
train_data <- cleaned_df

mlr_model <- training_mlr_model(train_data)
print(mlr_model)
```

```{r}
rf_model <- training_rf_model(train_data)
print(rf_model)
```
## 6. Model Testing
With `testing()`, you can test a saved model on a new dataset.

```{r, eval=FALSE}
test_results <- testing("models/mlr_model.rds", test_data)
print(test_results)
```

## 7. Visualization
To help interpretation, visualization functions help generate three types of plots: confusion matrix, random forest feature importance, and predicted probabilities.
```{r}
# 1. Confusion Matrix Plot
helper_conf_table <- data.frame(
  True = c("low risk", "mid risk", "high risk",
           "low risk", "mid risk", "high risk",
           "low risk", "mid risk", "high risk"),
  Predicted = c("low risk", "low risk", "low risk",
                "mid risk", "mid risk", "mid risk",
                "high risk", "high risk", "high risk"),
  Frequency = c(9, 17, 9, 10, 6, 9, 8, 14, 18),
  Percentage = c(33.3, 45.9, 25.0, 37.0, 16.2, 25.0, 29.6, 37.8, 50.0)
)

plot_conf_matrix(helper_conf_table, conf_type = "mlr", output_dir = tempdir())
```

```{r, eval=FALSE}
# 2. Random Forest Feature Importance Plot
# We avoid executing this because the example train_data has limited observations. 

rf_model <- training_rf_model(train_data)
plot_feature_importance(rf_model, output_dir = tempdir())
```

```{r}
# 3. Predicted Probability vs. Blood Sugar Plot

prob_table <- data.frame(
  BS = rep(seq(5.5, 8.0, by = 0.1), 3),
  Probability = runif(78, 0.2, 0.9),
  RiskLevel = factor(rep(c("low risk", "mid risk", "high risk"), each = 26))
)

plot_pred_prob(prob_table, output_dir = tempdir())
```




















